Setting up Data

### 2020-05-14 src/SetupData.Rmd
extensive debugging and organizing the datascripts

to Do

- try calq correlation of MLH1 counts across duplicates

- format the DF for applying to Mixed models

- try coding i) date image for each cell image, ii) most recent image date for each mouse

```{r setup, include=FALSE}
library(plyr)
library(dplyr)
library(raster)#for cV
library(ggplot2)
library(knitr)
library(kableExtra)
library(here)
library(xlsx)
#knitr::opts_chunk$set(echo = FALSE)

@@

@@

colors_of_strains <- c('WSB'= "#56B4E9",'G' ='cadetblue','LEW'= 'lightblue','PERC'= 'blue',
                       #blues

@@

```{r adj_nMLH1, echo=FALSE}
#add a column with male adjusted MLH1 values (+1 to all males)
MLH1_data$adj_nMLH1.foci <- ifelse(MLH1_data$sex=="male", MLH1_data$nMLH1.foci+1, MLH1_data$nMLH1.foci)
MLH1_data$adj_nMLH1.foci <- as.numeric(MLH1_data$adj_nMLH1.foci)
```

@@

#order the batches in MLH1 and original
```

```{r kable.table, echo=FALSE}
#table comparing number of images in quantification batches
#ToDo:  correct the case of all batches, better names
#set an order to batches?
original_batches <- as.data.frame(table(original_DF$Batch) ) #these look bad
colnames(original_batches) <- c("BatchNum", "orginal")

@@

```{r all.data.table, echo=FALSE}

@@

```{r mouse.table, echo=FALSE}
#Make a mouse level table

AP_mouse_table.MLH1_HQ <- ddply(MLH1_data[MLH1_data$quality < 5,], c("mouse"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(nMLH1.foci),
                        mean_co = as.numeric(format(round(  mean(nMLH1.foci, na.rm=TRUE), 3 ), nsmall=3) ),
                        cV = cv(nMLH1.foci, na.rm=TRUE),

@@

#source("~./MLH1repo/src/CommonFunc_MLH1repo.R")

AP_mouse_table.MLH1_HQ <- add_strain(AP_mouse_table.MLH1_HQ)
AP_mouse_table.MLH1_HQ <- add_subsp(AP_mouse_table.MLH1_HQ)
AP_mouse_table.MLH1_HQ <- add_sex(AP_mouse_table.MLH1_HQ)
AP_mouse_table.MLH1_HQ <- add_category(AP_mouse_table.MLH1_HQ)

#add DOB to this table?
mice.frm.meta <- AP_mouse_table.MLH1_HQ[as.character(AP_mouse_table.MLH1_HQ$mouse) %in% as.character(meta.data$mouse),]

@@

#subsp in metadata not fully correct
meta.data <- add_strain(meta.data)
meta.data <- add_subsp(meta.data)
meta.data <- add_sex(meta.data)
meta.data <- add_category(meta.data)

#during merging -- try to just merge a few cols: 
AP_mouse_table_w.Ages <- merge(AP_mouse_table.MLH1_HQ, meta.data)

#this should be transfering the image dataes!

#don't use by, that makes duplicate cols
# by.y = "mouse", by.x = "mouse", all = TRUE)
#what are the differences between these two

#remove mice without MLH1 obs
AP_mouse_table_w.Ages <- AP_mouse_table_w.Ages[!(is.na(AP_mouse_table_w.Ages$Nmice) | AP_mouse_table_w.Ages$Nmice == ""), ]

#remove cols with bad suffix
#AP_mouse_table_w.Ages <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
#                            subsp.y,subsp.x,
#                    category.y,category.x, strain.x, strain.y,
#                    mouse.1  )  )

#add all the cols back
AP_mouse_table_w.Ages <- add_strain(AP_mouse_table_w.Ages)
AP_mouse_table_w.Ages <- add_subsp(AP_mouse_table_w.Ages)
AP_mouse_table_w.Ages <- add_sex(AP_mouse_table_w.Ages)
AP_mouse_table_w.Ages <- add_category(AP_mouse_table_w.Ages)
AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)

#make sure they are all factors - this was breaking the kniting
#AP_mouse_table.HQ$mouse <- as.factor(AP_mouse_table.HQ$mouse)
#AP_mouse_table.HQ$sex <- as.factor(AP_mouse_table.HQ$sex)
#AP_mouse_table.HQ$strain <- as.factor(AP_mouse_table.HQ$strain)
#AP_mouse_table.HQ$subsp <-  as.factor(AP_mouse_table.HQ$subsp)

old.mice.df <- AP_mouse_table_w.Ages[(AP_mouse_table_w.Ages$age.weeks > 25),]
#remove the NAs
old.mice.df <- old.mice.df[!(is.na(old.mice.df$mouse) | old.mice.df$mouse==""), ]

#remove juvi mice via the list

#remove bad ages
#AP_mouse_table_w.Ages <- AP_mouse_table_w.Ages[( (AP_mouse_table_w.Ages$age.weeks < 25) &  (AP_mouse_table_w.Ages$age.days != 13) ) | #is.na(AP_mouse_table_w.Ages$age.weeks), ]
```

@@

#REMOVE EXTRA STRAINS
Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$subsp != "Cast",]
#remove musc strains
Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$strain != "AST",]
Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$strain != "TOM",]
#remove PERC, MOLF, CZECH, (no females)
Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$strain != "CZECH",]
Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$strain != "PERC",]

@@

```{r HetC.DF.4MM, echo=FALSE}
#this block has so many fing errors

#for the DF for use in the HetC Mixed Model
#Which Start with the AP_mouse_table

#variance is still chr
AP_mouse_table_w.Ages$var <- as.numeric(AP_mouse_table_w.Ages$var)

#use the merged DF to filter by age 
#AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
#AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
#AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x
#main file, make everything unordered factors

#setwd("~./MLH1repo/src/")

source(here("src","CommonFunc_MLH1repo.R") )
AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)
DF.HetC.MixedModel <- AP_mouse_table_w.Ages %>% filter(species == "M.musculus")

@@

#AP_mouse_table_w.Ages
#use the above table to pull the male mouse levels
mlh1male_mouse.av <-  AP_mouse_table_w.Ages %>% filter(sex== "male")

#add the Rec.group markers

#mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM$strain), 1, 
#                            ifelse(grepl("MSM", mouse.avs_4MM$strain), 1,
#                             ifelse(grepl("SKIVE", mouse.avs_4MM$strain), 1, 0)) )

@@

male_mouse_table_w.Ages <- AP_mouse_table_w.Ages[(AP_mouse_table_w.Ages$sex == "male") &(AP_mouse_table_w.Ages$category != "other"),]

all.ages_plot <- ggplot(AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$sex == "male",], aes(age.weeks))+geom_histogram(binwidth = 3)+facet_wrap(~category)+ggtitle("Male Age Distributions")
all.ages_plot

@@

```{r color.scheme}
#color scheme, #mouse colors
#site for colors, #http://sape.inf.usi.ch/quick-reference/ggplot2/colour
colors_of_strains <- c('WSB'= "#56B4E9",'G' ='cadetblue','LEW'= 'lightblue','PERC'= 'blue',
                       #blues

                       'PWD'= 'red2',
                       'SKIVE'= 'coral1','KAZ'= 'coral4', 'TOM'= 'indianred1','AST' = 'red4',
                       'CZECH'= 'indianred4',

#DMC1 Sections

@@

>>>>

```{r setup, include=FALSE}
library(plyr)
library(dplyr)
library(raster)#for cV
library(ggplot2)
library(knitr)
library(kableExtra)
library(here)
library(xlsx)
library(reshape2)
#knitr::opts_chunk$set(echo = FALSE)

@@

#total SC -- the one file should be all I need?
Total.SC_DF = read.xlsx2("~./MLH1repo/data/Total.SC_merge.MLH1.xlsx", header=TRUE, 1)

#Curated Bivdata - this excel file is too big to work with read.xlsx2
#Curated_BivData = read.xlsx2("~./MLH1repo/data/Curated_BivData.xlsx", header=TRUE, 1)

Curated_BivData <-  read.csv("~./MLH1repo/data/Curated_BivData.csv", header=TRUE)

@@

#site for colors, #http://sape.inf.usi.ch/quick-reference/ggplot2/colour
colors_of_strains <- c('WSB'= "#56B4E9",'G' ='cadetblue','LEW'= 'lightblue','PERC'= 'blue',
                       #blues

@@

```{r adj_nMLH1, echo=FALSE, eval=FALSE}
#add a column with male adjusted MLH1 values (+1 to all males)
MLH1_data$adj_nMLH1.foci <- ifelse(MLH1_data$sex=="male", MLH1_data$nMLH1.foci+1, MLH1_data$nMLH1.foci)
MLH1_data$adj_nMLH1.foci <- as.numeric(MLH1_data$adj_nMLH1.foci)

#adjusted MLH1 no longer used
```

@@

#order the batches in MLH1 and original
#apply this ordering to all other DF?

```{r Batch.Stats, echo=FALSE, eval=TRUE}
#plot / table comparing number of images in quantification batches
#basic visualization for number of images in batches

#ToDo:  correct the case of all batches, better names
#set an order to batches?
original_batches <- as.data.frame(table(original_DF$Batch) ) #these look bad
colnames(original_batches) <- c("BatchNum", "orginal")

@@

```{r all.data.table, echo=FALSE, eval=FALSE}

@@

#NA cause by CAST female
#this table isn't used much
```

```{r HQ.mouse.table, echo=FALSE}
#Make a mouse level table

#change name to Mouse.Table_HQ from #AP_mouse_table.MLH1_HQ
#

Mouse.Table_HQ <- ddply(MLH1_data[MLH1_data$quality < 5,], c("mouse"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(nMLH1.foci),
                        mean_co = as.numeric(format(round(  mean(nMLH1.foci, na.rm=TRUE), 3 ), nsmall=3) ),
                        cV = cv(nMLH1.foci, na.rm=TRUE),

@@

Mouse.Table_HQ <- add_strain(Mouse.Table_HQ)
Mouse.Table_HQ <- add_subsp(Mouse.Table_HQ)
Mouse.Table_HQ <- add_sex(Mouse.Table_HQ)
Mouse.Table_HQ <- add_category(Mouse.Table_HQ)
Mouse.Table_HQ <- add_species(Mouse.Table_HQ)

@@

#subsp in metadata not fully correct
meta.data <- add_strain(meta.data)
meta.data <- add_subsp(meta.data)
meta.data <- add_sex(meta.data)
meta.data <- add_category(meta.data)
meta.data <- add_species(meta.data)

#add DOB to this table?
mice.frm.meta <- Mouse.Table_HQ[as.character(Mouse.Table_HQ$mouse) %in% as.character(meta.data$mouse),]

Mouse.Table_HQ.OG <- Mouse.Table_HQ

Mouse.Table_HQ <- Mouse.Table_HQ.OG

#during merging -- try to just merge a few cols: 
Mouse.Table_HQ <- merge(Mouse.Table_HQ, meta.data)
#this removes ~28 mice

old.mice.df <- Mouse.Table_HQ[(Mouse.Table_HQ$age.weeks > 25),]
#remove the NAs
old.mice.df <- old.mice.df[!(is.na(old.mice.df$mouse) | old.mice.df$mouse==""), ]

Mouse.Table_HQ <- Mouse.Table_HQ[ ! Mouse.Table_HQ$mouse %in% old.mice.df, ]

#remove juvi mice via the list
#remove bad ages
#this is removing female

#this removes everything but males in these categories
#Mouse.Table_HQ <- Mouse.Table_HQ[ (Mouse.Table_HQ$sex == "female") | ( #(Mouse.Table_HQ$sex == "male") & (Mouse.Table_HQ$age.weeks < 25) & #(Mouse.Table_HQ$age.days != 13) ), ]

#i don't understand why so many WSB female are removed..
# | is.na(Mouse.Table_HQ$age.weeks) | (Mouse.Table_HQ$age.weeks == 0)

@@

#REMOVE EXTRA STRAINS -- why?
#Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$subsp != "Cast",]
#remove musc strains
#Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$strain != "AST",]
#Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$strain != "TOM",]
#remove PERC, MOLF, CZECH, (no females)
#Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$strain != "CZECH",]
#Q12_mouse_table_w.Ages <- Q12_mouse_table_w.Ages[Q12_mouse_table_w.Ages$strain != "PERC",]

@@

```{r HetC.DF.4MM, echo=FALSE}

#variance is still chr
Mouse.Table_HQ$var <- as.numeric(Mouse.Table_HQ$var)

DF.HetC.MixedModel <- Mouse.Table_HQ %>% filter(species == "M.musculus")

@@

#AP_mouse_table_w.Ages
#use the above table to pull the male mouse levels
mlh1male_mouse.av <-  Mouse.Table_HQ %>% filter(sex== "male")

#add the Rec.group markers
#mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM$strain), 1, 
#                            ifelse(grepl("MSM", mouse.avs_4MM$strain), 1,
#                             ifelse(grepl("SKIVE", mouse.avs_4MM$strain), 1, 0)) )

@@

male_mouse_table_w.Ages <- Mouse.Table_HQ[(Mouse.Table_HQ$sex == "male") &(Mouse.Table_HQ$category != "other"),]

all.ages_plot <- ggplot(Mouse.Table_HQ[Mouse.Table_HQ$sex == "male",], aes(age.weeks))+geom_histogram(binwidth = 3)+facet_wrap(~category)+ggtitle("Male Age Distributions")
all.ages_plot

@@

# DMC1 Section

@@

# BivData

for these full curated, add all the labels and save them as excel files in the BivData setup file.

full curated, reduced datasets

Are the strain levels plots needed?

Curated_BivData will be the master DF with most observations. DFs for mixed models and tests need to subsetted from this one.

```{r no.Foci.Pos, echo=FALSE}
#no posittion
#just.SC.length <- Curated_BivData[is.na(Curated_BivData$Foci1),] #NAs not registering
just.SC.length <- Curated_BivData[(Curated_BivData$Foci1 == ""),]

#table(just.SC.length$category)
```

Main files from this Script:

  1. **Curated_BIVDATA**: cleaned up version of the passing, curated Bivalent measures

  2.  **Foci_TRUE_Curated_Bivdata**: filterd version of the Curated_BivData file, but with only measures that have 'good' foci position information

  3. OG.Curate_FULL: original version of the Bivalent data (before 'failed' SC wiped away)

  4. **MELT_ABS**: make melt version of foci positions for abs measuers

  5. CELL level table (full all mice) (use this to make long biv dataset)

  6. Long Biv Dataset

  7. mouse table (full all mice)

  8. **Strain DFs** for Het models (summary strain DFs and tables?)

  9. **Mouse.Table_BivData_4MM** Mouse level table for HetC mixed models

  9. **male Subsp DFs** for Polymorphism models **Male.poly.Mouse.Table_BivData_4MM**

# Male specific BivData set

not sure if SC.Length.male_table is used for anything.

```{r male.sc.length.strain, echo=FALSE}

#these measure ect should be in the male - mouse 
#isolate all musc and male
Curated_BivData_male <- Curated_BivData[ (Curated_BivData$sex == "male") & (Curated_BivData$species == "M.musculus"),]

#make strain average table
SC.Length.male_table <- ddply(Curated_BivData_male, c("strain"), summarise,
                             mean_SC = mean(chromosomeLength),
                              n_SC = length(Obj.ID),

                   mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),
                 # CO2 = sum(hand.foci.count == 2, na.rm=TRUE),

                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),

                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),

                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)

)

# Cell Level Table

```{r cell.level.table, echo=FALSE}

cell.level.BivData_table <- ddply(.data=Curated_BivData, 
                       .(fileName),
                       summarize, 
                       boxs.IDd = max(boxNumber),
                       nboxes_passed = length(boxNumber),

                       Nbiv = length(Obj.ID), #if there are na's below -- it breaks
                       CO0 =  sum(hand.foci.count == 0, na.rm=TRUE), 
                       CO1 =  sum(hand.foci.count == 1, na.rm=TRUE), 
                       CO2 = sum(hand.foci.count == 2, na.rm=TRUE),
                       CO3 = sum(hand.foci.count == 3, na.rm=TRUE),
                       CO4 = sum(hand.foci.count == 4, na.rm=TRUE),

                       ncells = length(unique(fileName)),
                       nbivs = length(unique(Obj.ID)),

                       mean.SC_length = mean(chromosomeLength, na.rm = TRUE),
                       var.SC_length = var(chromosomeLength),
                       sd.SC_length   = round(sd(chromosomeLength), 3),
                       se.SC_length   = round(sd.SC_length / sqrt(nbivs), 3), #check that nBivs is the right number

                       cv.SC = cv(chromosomeLength),

#AutoSC = sum( SC.length[which(blobjectClass != "XY")] ),
                      mean.IFD_all = mean(   IFD1_ABS, na.rm = TRUE),

                #this below might not be correct
                     mean.IFD_2C0 = mean(   IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),

                       var.IFD_2CO = var( IFD1_ABS[which(hand.foci.count == 2)], na.rm=TRUE),
                       #sd.IFD_2CO   = round(sd(IFD1, na.rm=TRUE), 3),
                       #se.IFD_2CO   = round(sd.IFD / sqrt(nbivs), 3)

#cent_dis
                      mean_cent_dist = mean(dis.cent, na.rm=TRUE),
                       Qar.4 = quantile(chromosomeLength, na.rm = TRUE)[[4]], #calq threshold for all cells, pull out 'almost.whole.cells' later
                      Qar.3 = quantile(chromosomeLength, na.rm = TRUE)[[3]],
                      Qar.2 = quantile(chromosomeLength, na.rm = TRUE)[[2]],
                      Qar.1 = quantile(chromosomeLength, na.rm = TRUE)[[1]]

                        )

cell.level.BivData_table <- add_mouse(cell.level.BivData_table)
cell.level.BivData_table <- add_sex(cell.level.BivData_table)
cell.level.BivData_table <- add_strain(cell.level.BivData_table)
cell.level.BivData_table <- add_category(cell.level.BivData_table)
cell.level.BivData_table <- add_species(cell.level.BivData_table)

#as.data.frame( head(cell.level.BivData_table) )
#use quartille 2 as a min, on average it grabs 4 
```

```{r show.pass.rate, echo=FALSE}

pass.rate.hist <- ggplot(cell.level.BivData_table, aes(x = nboxes_passed) )+geom_histogram(binwidth = 1, alpha=.6) + facet_wrap(~category)+ggtitle("number of objects ID'd per cell")
pass.rate.hist

make histogram of the pass rate per cell.

3 cells have manual measures and BivData measures. test error, need to match them up. There are 145 cells from the BivData set which are close to whole cells.  
8jun15_23apr15_PWD_f2_sp1_20_rev, 29nov16_13nov16_MSM_m2_sp1_13_rev, 27oct16_31aug16_MSM_m1_sp1_9_rev

# Reduced Bivalent Dataset (Long and Short)

I want to make a dataset of 'longest' 25% chromosomes from cells with at least 16 segmented bivalents from a single cell.

```{r Reduced.Biv.DataSet.chrm.dataset, echo=FALSE}

#1. Isolate bivalent measrues from 'almost whole cells'   (how many biv's are from cells with >15 bivalents)
over_15_biv_cell_DF <- cell.level.BivData_table[cell.level.BivData_table$Nbiv >15, ]
BivData_from_15biv_cells <- Curated_BivData[ Curated_BivData$fileName %in% over_15_biv_cell_DF$fileName, ]

#check for blank obj.id
BivData_from_15biv_cells$Obj.ID <- paste(BivData_from_15biv_cells$fileName, BivData_from_15biv_cells$boxNumber, sep = "_")

#check for dup
#length(duplicated(BivData_from_15biv_cells$Obj.ID))#2605
#anyDuplicated(BivData_from_15biv_cells$Obj.ID)#1997
#dups.dummy <- BivData_from_15biv_cells[duplicated(BivData_from_15biv_cells$Obj.ID),]
#no duplicates

#remove duplicate obj.ids
#BivData_from_15biv_cells <- BivData_from_15biv_cells %>% distinct() #distinct doesn't work for chr type
#BivData_from_15biv_cells <- BivData_from_15biv_cells[ ! BivData_from_15biv_cells$fileName %in% dups.dummy$fileName, ]

#make sure obj.id and fileName are charcter for below
BivData_from_15biv_cells$fileName <- as.character(BivData_from_15biv_cells$fileName)
BivData_from_15biv_cells$Obj.ID <- as.character(BivData_from_15biv_cells$Obj.ID)

missing_from_cell_table <- BivData_from_15biv_cells[!BivData_from_15biv_cells$fileName %in% cell.level.BivData_table$fileName, ] 
#all these files stem from

#these are bivalents which don't appear the in cell level table, which can break the loop
BivData_from_15biv_cells <- BivData_from_15biv_cells[ ! BivData_from_15biv_cells$fileName %in% missing_from_cell_table$fileName, ]
#final cleaned BivData table, 2428

#2. calq the Q4 threshold
#loop for greping the mouse from file name ect
#each Q is a quartille

#mini.now  <- BivData_from_15biv_cells[1:300,]

for(  i in 1:length( unique(BivData_from_15biv_cells$Obj.ID) ) ){

  newfile.str = strsplit( BivData_from_15biv_cells$Obj.ID[i], "_" )[[1]][1:7]
  new.egg <- paste0(newfile.str, collapse = "_")

  the.Q4 <- cell.level.BivData_table$Qar.4[ which( grepl(new.egg, cell.level.BivData_table$fileName ) )]

  the.Q3 <-cell.level.BivData_table$Qar.3[ which( grepl(new.egg, cell.level.BivData_table$fileName ) )]

  the.Q2 <-cell.level.BivData_table$Qar.2[ which( grepl(new.egg, cell.level.BivData_table$fileName ) )]

  the.Q1 <-cell.level.BivData_table$Qar.1[ which( grepl(new.egg, cell.level.BivData_table$fileName ) )]

  BivData_from_15biv_cells$cell.Q4[i] <- the.Q4
    BivData_from_15biv_cells$cell.Q3[i] <- the.Q3
  BivData_from_15biv_cells$cell.Q2[i] <- the.Q2
  BivData_from_15biv_cells$cell.Q1[i] <- the.Q1

  the.FN <- as.character(cell.level.BivData_table$fileName[which(grepl(new.egg, cell.level.BivData_table$fileName ) )] )

  BivData_from_15biv_cells$Q4.FN[i] <- as.character(the.FN)
  BivData_from_15biv_cells$Q1.FN[i] <- as.character(the.FN)
}

#3.mark which bivalents are above the threshold

BivData_from_15biv_cells$long.Biv <- ifelse( (BivData_from_15biv_cells$chromosomeLength >= as.numeric(BivData_from_15biv_cells$cell.Q4)  ), "yes", "no" )

real.long.bivData <- BivData_from_15biv_cells[BivData_from_15biv_cells$long.Biv == "yes",]#713 observations

#isolate the short bivdata
BivData_from_15biv_cells$short.Biv <- ifelse( (BivData_from_15biv_cells$chromosomeLength <= as.numeric(BivData_from_15biv_cells$cell.Q2)  ), "yes", "no" )

real.short.bivData <- BivData_from_15biv_cells[BivData_from_15biv_cells$short.Biv == "yes",]# 708 much few than

```{r Look.short.biv.data, echo=FALSE, include=FALSE, eval=FALSE}

#showing SC lengths per cell, with lines for the quartille values.

mini_G.15_15biv_FN <- BivData_from_15biv_cells %>% filter(category == "G male")
#ploting short biv data
G.15_15biv_male <- BivData_from_15biv_cells %>% filter(category == "G male")
mini_G.15_15biv_FN <- sample( G.15_15biv_male$fileName, 10)
mini_G.15_15biv  <- BivData_from_15biv_cells[BivData_from_15biv_cells$fileName %in% mini_G.15_15biv_FN, ]

ppll <- ggplot(mini_G.15_15biv, aes(x = fileName, y=chromosomeLength, color=mouse))+geom_jitter()+facet_wrap(~fileName, scales="free")+ theme(legend.position="none")+
  geom_hline(aes(yintercept= cell.Q1), color='coral') + 
  geom_hline(aes(yintercept= cell.Q2), color='red') +
 geom_hline(aes(yintercept= cell.Q3), color='blue') +   ##actually Q3 might not be a real thing...
 geom_hline(aes(yintercept= cell.Q4), color='black')

  #this plot shows that the Q1 is capturing only single values -- I should try Q2
#use Q2 as a 
```

```{r table.long.biv, echo=FALSE}

# make a table for the Mixed Models -- from the long.BivData
# Calq mouse averages

#add in all the other stats there are in the other tables

Long.biv_mouse_table <- ddply(real.long.bivData, c("mouse"), summarise,

                        Nmice = length(unique(mouse)),
                        Ncells  = length(Obj.ID),

                      mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
                        cV = cv(chromosomeLength),
                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
                        sd   = round(sd(chromosomeLength), 3),
                        se   = round(sd / sqrt(Ncells), 3),

                        #SC lengths by class
                        nCO0 =  sum(hand.foci.count == 0, na.rm=TRUE), 
                      nCO1 =  sum(hand.foci.count == 1, na.rm=TRUE), 
                       nCO2 = sum(hand.foci.count == 2, na.rm=TRUE),
                       nCO3 = sum(hand.foci.count == 3, na.rm=TRUE),
                       nCO4 = sum(hand.foci.count == 4, na.rm=TRUE),

                      mean.siscoten_0CO = mean(SisCoTen[which(hand.foci.count == 0)], na.rm = TRUE),
                      mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
                     mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
                      mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),

                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),

                 mean_IFD.2CO_ABS = mean(IFD1_ABS[which(hand.foci.count == 2)], na.rm = TRUE),
                 mean_IFD.2CO_PER = mean(IFD1_PER[which(hand.foci.count == 2)], na.rm = TRUE),

                       #CO position things, requires the melt?
              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)])

        #          mean_siscoten = mean(siscoten)
        #
)
#mouse averages for long.biv

Long.biv_mouse_table <- add_sex(Long.biv_mouse_table)
Long.biv_mouse_table <- add_strain(Long.biv_mouse_table)
Long.biv_mouse_table <- add_subsp(Long.biv_mouse_table)
Long.biv_mouse_table <- add_category(Long.biv_mouse_table)

#unordered for the mixed models (move this to the MM blocks...?)
Long.biv_mouse_table$subsp <- factor(Long.biv_mouse_table$subsp, ordered = FALSE)
Long.biv_mouse_table$sex <- factor(Long.biv_mouse_table$sex, ordered = FALSE)
Long.biv_mouse_table$strain <- factor(Long.biv_mouse_table$strain, ordered = FALSE)

#rec group
Long.biv_mouse_table$Rec.group <- ifelse(grepl("PWD", Long.biv_mouse_table$category), 1, 
                            ifelse(grepl("MSM", Long.biv_mouse_table$category), 1,
                             ifelse(grepl("SKIVE", Long.biv_mouse_table$category), 1, 0)) )

#long.biv. 1) HetC  2) male.poly (the main factors are still unordered)

#1) Long Biv for 
#remove all the non-used strains!!
Long.biv_mouse_table_HetC <- Long.biv_mouse_table

Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[Long.biv_mouse_table$strain != "AST",]
Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[Curated_BivData$strain != "TOM",]
Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[Long.biv_mouse_table_HetC$strain != "CZECH",]
Long.biv_mouse_table_HetC$strain <- droplevels(Long.biv_mouse_table_HetC$strain)

#remove na's
Long.biv_mouse_table_HetC <- Long.biv_mouse_table_HetC[!(is.na(Long.biv_mouse_table_HetC$mouse) | Long.biv_mouse_table_HetC$mouse==""), ]

#table(droplevels( Long.biv_mouse_table_HetC$category) )  #for double checking

#2) Long Biv for male.poly
#just males, just Musculus
Long.biv_mouse_table_male.poly <- Long.biv_mouse_table

Long.biv_mouse_table_male.poly <- Long.biv_mouse_table_male.poly[Long.biv_mouse_table_male.poly$sex == "male",]

#table(droplevels( Long.biv_mouse_table_male.poly$category) ) #for double checking
```

The above code chunk makes 2 mouse level tables; **Long.biv_mouse_table_HetC** and **Long.biv_mouse_table_male.poly**. Run these through the model systems.

# Mouse Level Single BivData Table

Mouse level tables are the main meat of the models. There will be several for use in different models. Format these tables for run the mixed models (unordered factors, ect).

The block below creates clean.BivData, which removes ~600 rows if just SC measures
(This might be important for the proportion code)

```{r FULL.mouse.BivData.table, echo=FALSE}

#length( Curated_BivData %>% filter(hand.foci.count == "NA") )#36

table(Curated_BivData$hand.foci.count, exclude = NULL)
#there are ~600 rows with SC and no hand foci count (blah)

#remove bivalent entries without hand foci count values
cleanCurated_BivData <-  Curated_BivData  %>%  filter(hand.foci.count  == 0 | hand.foci.count  == 1 |
                                      hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 )

mouse.level.BivData_table <- ddply(.data=cleanCurated_BivData, 
                       .(mouse),
                       summarize,

                       ncells = length(unique(fileName)),

                       boxs.IDd = max(boxNumber),
                       nboxes_passed = length(boxNumber),

                       Nbiv = length(Obj.ID), #if there are na's below -- it breaks

                      tot.biv = length( !is.na(hand.foci.count) ),

                       n0CO = sum( (hand.foci.count == 0), na.rm = TRUE ),
                      n1CO = sum( (hand.foci.count == 1), na.rm = TRUE ),
                      n2CO = sum( (hand.foci.count == 2), na.rm = TRUE ),
                      n3CO = sum( (hand.foci.count == 3), na.rm = TRUE ),
                      n4CO = sum(hand.foci.count == 4, na.rm=TRUE),

                       mean.SC_length = mean(chromosomeLength, na.rm = TRUE),
                       var.SC_length = var(chromosomeLength, na.rm = TRUE),
                       sd.SC_length   = round(sd(chromosomeLength), 3),
                       se.SC_length   = round(sd.SC_length / sqrt(Nbiv), 3),

                       #sis.co.ten
                       mean.sis.co.ten = mean(SisCoTen, na.rm = TRUE)
)

mouse.level.BivData_table <- add_sex(mouse.level.BivData_table)
mouse.level.BivData_table <- add_strain(mouse.level.BivData_table)
mouse.level.BivData_table <- add_subsp(mouse.level.BivData_table)
mouse.level.BivData_table <- add_category(mouse.level.BivData_table)

#unordered for the mixed models (move this to the MM blocks...?)
mouse.level.BivData_table$subsp <- factor(mouse.level.BivData_table$subsp, ordered = FALSE)
mouse.level.BivData_table$sex <- factor(mouse.level.BivData_table$sex, ordered = FALSE)
mouse.level.BivData_table$strain <- factor(mouse.level.BivData_table$strain, ordered = FALSE)

mouse.level.BivData_table$Rec.group <- ifelse(grepl("PWD male", mouse.level.BivData_table$category), 1, 
                              ifelse(grepl("MSM male", mouse.level.BivData_table$category), 1,
                                 ifelse(grepl("SKIVE male", mouse.level.BivData_table$category), 1, 0)))
```

```{r mouse.table.HetC.MM, echo=FALSE}

#this mouse level table -- should not have strains without female obs
#only Mmusculus

Mouse.Table_BivData_4MM <- ddply(Curated_BivData, c("mouse"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(  unique(fileName) ),
                      tot.biv = length(Obj.ID),

                      n0CO = sum( (hand.foci.count == 0), na.rm = TRUE ),
                      n1CO = sum( (hand.foci.count == 1), na.rm = TRUE ),
                      n2CO = sum( (hand.foci.count == 2), na.rm = TRUE ),
                      n3CO = sum( (hand.foci.count == 3), na.rm = TRUE ),

                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
                        cV = cv(chromosomeLength),
                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
                        sd   = round(sd(chromosomeLength), 3),
                        se   = round(sd / sqrt(Ncells), 3),

                       #calq mean.sc by chrm class
                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),
